Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-01 main.tex
spell checker

\cye{Our main purpose here is to fill this gap. In the gradient case, we can go over with so-called face-connected triangulations. However, to proceed} in the \cye{general exterior derivative case, we are brought to} so-called shellable triangulations. 
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important stepping stone towards more general cases. \cye{Still, we cover most practical nonconvex and non-star-shaped triangulated domains, and as a particular example, local triangulations} around a simplex within a larger triangulation \cye{(the so-called local patches or stars), which are} shellable in two and three spaces dimensions. \cye{Over convex bounded domains, we derive Poincar\'e--Friedrichs inequalities with computable constants for the whole $L^{p}$ de~Rham complex.}

@@

Not much attention\cye{, however,} seems to have been given to the study of \cye{computable} constants in~\eqref{math:pot_PF_curl_div}, even if the domain $\Omega$ is convex. \cye{Important results in the Hilbert case $p=2$ can be found in~\cite{guerini2004eigenvalue, Paul_Vald_PF_grad_curl_div_20} that we discuss now in the broder setting of} the exterior calculus formalism\cye{, but the Banach case $p \neq 2$ seems largely unexplored.}

@@

Exterior calculus~\cite{greub1967multilinear,lee2012smooth} is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite, gross2004electromagnetic, arnold2006finite, arnold2009geometric, arnold2010finite, demlow2014posteriori, licht2021local, arnold2021complexes}. 
This formalism is independent of the \cye{space} dimension and highlights \cye{extends the} underlying geometric structures common to the gradient, curl, and divergence operators \cye{in three space dimensions}.
\cye{In this setting, the common writing of the Poincar\'e--Friedrichs inequalities we are interested in is just as~\eqref{math:PF_gradient_1}: for every $w \in W^{p}\Alt^{k}(\Omega)$, there exists $u \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ and such that (all the notation is in detail fixed in Section~\ref{section:calculus} below)
\begin{align} \label{math:PF_DF} 
    \| u \|_{L^{p}(\Omega} \leq  C_{k,\Omega,p} \| \cartan w \|_{L^{p}(\Omega)}.
\end{align} 
}\mwl{MV: The roles of $u$ and $w$ are flipped in the various parts of the paper, see~\eqref{math:PF_DF} in comparsion with Lemma~\ref{lemma:PFexteriorsimplex}. It would be great to unify.}

@@

\cye{The constant in~\eqref{math:pot_PF_grad} for $p=2$ corresponds to a lower bound for the Laplace eigenvalues and quantifies the stability properties of the Laplacian on the domain $\Omega$.} 
\cye{Similarly, t}he constant \cye{in~\eqref{math:pot_PF_curl} for $p=2$} corresponds to a lower bound for the Maxwell eigenvalues and quantifies the stability properties of the Maxwell system on $\Omega$. 
Thus, computable upper bounds on the Poincar\'e--Friedrichs constants also give computable lower bounds for the eigenvalues of the associated Laplacians \cye{and vice versa}. 
\cye{This eigenvalues optics, upon employing a suitable numerical simulation} over a general triangulation \cye{and some clever postprocessing estimates, enables to numerically compute guaranteed upper bounds on the Poincar\'e--Friedrichs constants, see~\cite{Cars_Ged_LB_eigs_14,Liu_fram_eigs_15} for~\eqref{math:pot_PF_grad} and~\cite{gallistl2023computational} for \eqref{math:pot_PF_curl}}. \cye{Our approaches, in contrast, are either analytical or combinatorial, and in any case we only employ the given geometry but no computer simulation.}

@@

\cye{We devote a considerable effort to convex domains for two reasons: first, they are evidently of independent interest, and second, we need them as a building block for the above-discussed recursive procedures on general (nonconvex and non-star-shaped) triangulated domains. Here,} we follow closely the \cye{concepts of} Costabel and McIntosh~\cite{costabel2010bogovskiui} who study regularized Poincar\'e and Bogovski\u{\i}-type potential operators. In comparison to their exposition, \cye{which allows for} domains star-shaped with respect to a ball, the regularized potential operators we propose are simplified: we only study them over convex domains and use simpler (constant) weight functions. 
The resulting operator features generally lower regularity \cye{but} its properties are sufficient for our purposes. \cye{Crucially, this allows us to} easily estimate their operator norms.

@@

\cye{We treat specifically the gradient operator (the setting of Section~\ref{section:intro_grad}) on triangulated domains that can be nonconvex and non-star-shaped but where the triangulations are face connected (though not necessarily shellable). Here, Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} are our main results.} 
\mwl{MV: remark from after Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} moved here.} 
The computable Poincar\'e--Friedrichs constants obtained here depend on only a few parameters of the \cye{given} triangulation: the length of any traversal from the root simplex, the ratios of the volumes of \cye{the} adjacent simplices, the Poincar\'e--Friedrichs constants on each face patch, and the Poincar\'e--Friedrichs constant of the initial simplex. \cye{Here,} the \cye{face patch} Poincar\'e--Friedrichs constants \cye{are} estimated in terms of shape \cye{regularity parameters of the triangulation}; if the face patches of the triangulation are convex, then better estimates are possible.

\cye{Our main result for the curl and divergence operators (Section~\ref{section:intro_curl_div}) and more generally in the exterior calculus setting (Section~\ref{section:intro_EC}) for triangulated domains are Theorems~\ref{theorem:poincarefriedrichsestimate:exterior} and~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.
They treat triangulated domains that can be nonconvex and non-star-shaped but where the triangulation is shellable. As discussed above, though this is slightly less generic than merely face-connected triangulations, this namely covers patches of simplices sharing a given vertex, edge, face, or $n$-simplex (stars) in two and three space dimensions, since these local patches are shellable. The dependencies of the obtained computable Poincar\'e--Friedrichs constants are similar to those of the gradient operator in Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei}.}

@@

    Without loss of generality, $F$ lies in the hyerplane plane of the second to last coordinates,
    so that the height of $T_1$ and $T_2$ is in the first coordinate direction. 
    After further rotation, we can assume that $z_1$ and $z_2$ are in the span of the first three coordinate directions with 
    We can write 
    \begin{align*}
        z_1 = a_1 e_1 + a_2 e_2, \qquad z_2 = b_1 e_1 + b_2 e_2 + e_3 b_3.
    \end{align*}
    The face $F$ spans a hyperplane over which $\Xi$ is the identity, and it remains to study the matrix
    \begin{align*}
        A = 
        \begin{pmatrix}
            b_1/a_1 & 0 & 0
            \\
            (b_2-a_2)/a_1 & 1 & 0
            \\
            b_3/a_1 & 0 & 1
        \end{pmatrix}
        =:
        \begin{pmatrix}
            c_1 & 0 & 0
            \\
            c_2 & 1 & 0
            \\
            c_3 & 0 & 1
        \end{pmatrix}
    \end{align*}
    Its singular values are $1$ and 
    \begin{align*}
        \sigma_{\max}
        =
        \frac{1}{\sqrt 2}
        \sqrt{
            1 + c_1^2 + c_2^2 + c_3^2 
            +
            \sqrt{ ( 1 + c_1^2 + c_2^2 + c_3^2 )^2 - 4 c_1^2 }
        }
        =
        \frac{1}{\sqrt 2}
        \sqrt{
            1 + c_1^2 + c_2^2 + c_3^2 
            +
            \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 + 2 c_1 }
            \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 - 2 c_1 }
        }
        =
        \frac{1}{2}
        \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 + 2 c_1 }
        +
        \frac{1}{2}
        \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 - 2 c_1 }
        =
        \frac{1}{2}
        \sqrt{ ( 1 + c_1 )^2 + c_2^2 + c_3^2 }
        +
        \frac{1}{2}
        \sqrt{ ( 1 - c_1 )^2 + c_2^2 + c_3^2 }
        \leq 
        \sqrt{ ( 1 + c_1 )^2 + c_2^2 + c_3^2 }
        .
    \end{align*}
    \begin{align*}
        \sigma_{\max}
        =
        \frac{1}{\sqrt 2}
        \sqrt{
            1 + c_1^2 + c_2^2 + c_3^2 
            -
            \sqrt{ c_1^2 + c_2^2 + c_3^2 - 4 c_1^2 }
        }
        =
        \frac{1}{2}
        \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 + 2 c_1 }
        -
        \frac{1}{2}
        \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 - 2 c_1 }
        .
    \end{align*}

@@

\begin{remark}
    % The main feature in proving these results is that the best constant in the Poincar\'e inequality is already the constant in the $1$-dimensional case. 
    Any estimate for the Poincar\'e--Friedrichs constant implies an estimate for the Poincar\'e constant via~\eqref{math:p_vs_pf}. 
    Let us compare the estimate $C_{{\rm P},{\rm CW},p}$ for the Poincar\'e constant
    with the estimate obtained from the Poincar\'e--Friedrichs constant $C_{\PF,{\rm EFNT},p}$. 
    In the case $2 \leq p$, 
    \begin{align*}
        2^{1-\frac 2 p}
        C_{\PF,{\rm EFNT},p}
        =
        \frac{ 2^{1-\frac 2 p} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq 
        4^{-\frac 1 p}
        \leq
        C_{{\rm P},{\rm CW},p} 
        .
    \end{align*}
    In the case $2 \geq p$, 
    \begin{align*}
        2^{\frac 2 p - 1}
        C_{\PF,{\rm EFNT},p}
        =
        \frac{ 2^{\frac 2 p - 1} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq
        \frac{ 2^{\frac 2 p-1} }{2}
        &
        =
        2^{\frac 2 p - 2}
        \\&
        =
        4^{\frac 1 p - 1}
        \leq
        C_{{\rm P},{\rm CW},p} 
%         % 
%         .
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 2\pi \sqrt[p]{p-1} \cdot 2 }
%         =
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 4\pi \sqrt[p]{p-1} }
%         \leq 
%         \frac{1}{2}
        .
    \end{align*}
    It follows that \eqref{math:esposito:poincare} is generally a tighter estimate than \eqref{math:chuawheeden} for $1 < p < \infty$.
    % DONE: what exactly is the limit as p -> 1 ? It converges to 1.
    % DONT: insert the original inequality by chua and Wheeden
    % DONE: chua wheeden upper bound is not optimal!
    % DONE: Theorem 1.2 in Chua Wheeden 
% DONT: Keep ?
%     In the limit as $p$ goes to $1$, the factor $C_{\PF,{\rm EFNT},p}$ converges to $1/2$,
%     which is the optimal Poincar\'e--Friedrichs constant and also the optimal factor in the Poincar\'e inequality \eqref{math:acostaduran}
%     \begin{align*}
%         \| u - u_{\Omega} \|_{L^{1}(\Omega)}
%         \leq 
%         \frac{1}{2}
%         \diam(\Omega)
%         \| \nabla u \|_{L^{1}(\Omega)}
%         ,
%         \quad
%         \forall 
%         u \in W^{1,1}(\Omega)
%         .
%     \end{align*}
%     In the limit as $p$ goes to infinity, the factor $C_{\PF,{\rm EFNT},p}$ converges to $1/2$, 
%     which implies the Poincar\'e inequality 
%     \begin{align*}
%         \| u - u_{\Omega} \|_{L^{\infty}(\Omega)}
%         \leq 
%         \diam(\Omega)
%         \| \nabla u \|_{L^{\infty}(\Omega)}
%         ,
%         \quad
%         \forall 
%         u \in W^{1,\infty}(\Omega)
%         %
%     \end{align*}
%     because the $L^{\infty}$-norm is the limit of the $L^{p}$-norm. % DONT: cite 
%     % DONT: review this entire result 
\end{remark}

@@

\begin{remark}
    % The main feature in proving these results is that the best constant in the Poincar\'e inequality is already the constant in the $1$-dimensional case. 
    Any estimate for the Poincar\'e--Friedrichs constant implies an estimate for the Poincar\'e constant, via~\eqref{math:p_vs_pf}. 
    Let us compare the estimate $C_{{\rm P},{\rm CW},p}$ for the Poincar\'e constant
    with the estimate obtained from the Poincar\'e--Friedrichs constant $C_{\PF,{\rm EFNT},p}$. 
    In the case $2 \leq p$, 
    \begin{align*}
        2^{1-\frac 2 p}
        C_{\PF,{\rm EFNT},p}
        =
        \frac{ 2^{1-\frac 2 p} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq 
        4^{-\frac 1 p}
        \leq
        C_{{\rm P},{\rm CW},p} 
        .
    \end{align*}
    In the case $2 \geq p$, 
    \begin{align*}
        2^{\frac 2 p - 1}
        C_{\PF,{\rm EFNT},p}
        =
        \frac{ 2^{\frac 2 p - 1} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq
        \frac{ 2^{\frac 2 p-1} }{2}
        &
        =
        2^{\frac 2 p - 2}
        \\&
        =
        4^{\frac 1 p - 1}
        \leq
        C_{{\rm P},{\rm CW},p} 
%         % 
%         .
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 2\pi \sqrt[p]{p-1} \cdot 2 }
%         =
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 4\pi \sqrt[p]{p-1} }
%         \leq 
%         \frac{1}{2}
        .
    \end{align*}
    It follows that \eqref{math:esposito:poincare} is generally a tighter estimate than \eqref{math:chuawheeden} for $1 < p < \infty$.
    % DONE: what exactly is the limit as p -> 1 ? It converges to 1.
    % DONT: insert the original inequality by chua and Wheeden
    % DONE: chua wheeden upper bound is not optimal!
    % DONE: Theorem 1.2 in Chua Wheeden 
% DONT: Keep ?
%     In the limit as $p$ goes to $1$, the factor $C_{\PF,{\rm EFNT},p}$ converges to $1/2$,
%     which is the optimal Poincar\'e--Friedrichs constant and also the optimal factor in the Poincar\'e inequality \eqref{math:acostaduran}
%     \begin{align*}
%         \| u - u_{\Omega} \|_{L^{1}(\Omega)}
%         \leq 
%         \frac{1}{2}
%         \diam(\Omega)
%         \| \nabla u \|_{L^{1}(\Omega)}
%         ,
%         \quad
%         \forall 
%         u \in W^{1,1}(\Omega)
%         .
%     \end{align*}
%     In the limit as $p$ goes to infinity, the factor $C_{\PF,{\rm EFNT},p}$ converges to $1/2$, 
%     which implies the Poincar\'e inequality 
%     \begin{align*}
%         \| u - u_{\Omega} \|_{L^{\infty}(\Omega)}
%         \leq 
%         \diam(\Omega)
%         \| \nabla u \|_{L^{\infty}(\Omega)}
%         ,
%         \quad
%         \forall 
%         u \in W^{1,\infty}(\Omega)
%         %
%     \end{align*}
%     because the $L^{\infty}$-norm is the limit of the $L^{p}$-norm. % DONT: cite 
%     % DONT: review this entire result 
\end{remark}

@@

% While the underlying idea remains the same, they differ in the way the intermdiate potentials are extended.
% This is described in Theorem~\ref{theorem:poincarefriedrichsestimate:grad} and relies on Lemma~\ref{lemma:poincarefriedrichsoverfacepatch} as an auxiliary result. 
% This is described in Theorem~\ref{theorem:poincarefriedrichsestimate:gradzwei}.

@@

\begin{remark}
    The Poincar\'e--Friedrichs constant obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad} increasingly overestimates the best one as number of $n$-simplices in the triangulation $\calT$ increases. We hence target its use to local patches (stars), in particular nonconvex boundary stars. The latter occur inevitably at reentrant corners. Clearly, the same building principle applies whenever we have any nonoverlapping partition of $\overline \Omega$ into convex local patches of $n$-simplices $\{ U_m \}$ with a face-connected path. The proof proceeds verbatim, where we merely replace the simplices $\{ T_{m} \}$ by the convex local patches $\{ U_m \}$. This may allow to a large decrease of the number of elements in the partition of $\overline \Omega$ and thus to a largely improved estimate of the best Poincar\'e--Friedrichs constant. 
\end{remark}

@@

We now collect important examples of shellable triangulations. Essentially, in two space dimensions, interesting triangulations shellable, but starting from three space dimensions, nonshellable situations can arise. Our main interest are local patches (stars) within triangulations: these are shellable up to three space dimensions, but not necessarily beyond.

>>>>

\cye{Our main purpose here is to fill this gap. In the gradient case, we can go over with so-called face-connected triangulations. However, to proceed} in the \cye{general exterior derivative case, we are brought to} so-called shellable triangulations. 
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important stepping stone towards more general cases. \cye{Still, we cover most practical non-convex and non-star-shaped triangulated domains, and as a particular example, local triangulations} around a simplex within a larger triangulation \cye{(the so-called local patches or stars), which are} shellable in two and three spaces dimensions. \cye{Over convex bounded domains, we derive Poincar\'e--Friedrichs inequalities with computable constants for the whole $L^{p}$ de~Rham complex.}

@@

Not much attention\cye{, however,} seems to have been given to the study of \cye{computable} constants in~\eqref{math:pot_PF_curl_div}, even if the domain $\Omega$ is convex. \cye{Important results in the Hilbert case $p=2$ can be found in~\cite{guerini2004eigenvalue, Paul_Vald_PF_grad_curl_div_20} that we discuss now in the broader setting of} the exterior calculus formalism\cye{, but the Banach case $p \neq 2$ seems largely unexplored.}

@@

Exterior calculus~\cite{greub1967multilinear,lee2012smooth} is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite, gross2004electromagnetic, arnold2006finite, arnold2009geometric, arnold2010finite, demlow2014posteriori, licht2021local, arnold2021complexes}. 
This formalism is independent of the \cye{space} dimension and highlights \cye{extends the} underlying geometric structures common to the gradient, curl, and divergence operators \cye{in three space dimensions}.
\cye{In this setting, the common writing of the Poincar\'e--Friedrichs inequalities we are interested in is just as~\eqref{math:PF_gradient_1}: for every $w \in W^{p}\Alt^{k}(\Omega)$, there exists $u \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ and such that (all the notation is in detail fixed in Section~\ref{section:calculus} below)
\begin{align} \label{math:PF_DF} 
    \| u \|_{L^{p}(\Omega} \leq  C_{k,\Omega,p} \| \cartan w \|_{L^{p}(\Omega)}.
\end{align} 
}\mwl{MV: The roles of $u$ and $w$ are flipped in the various parts of the paper, see~\eqref{math:PF_DF} in comparison with Lemma~\ref{lemma:PFexteriorsimplex}. It would be great to unify.}

@@

\cye{The constant in~\eqref{math:pot_PF_grad} for $p=2$ corresponds to a lower bound for the Laplace eigenvalues and quantifies the stability properties of the Laplacian on the domain $\Omega$.} 
\cye{Similarly, t}he constant \cye{in~\eqref{math:pot_PF_curl} for $p=2$} corresponds to a lower bound for the Maxwell eigenvalues and quantifies the stability properties of the Maxwell system on $\Omega$. 
Thus, computable upper bounds on the Poincar\'e--Friedrichs constants also give computable lower bounds for the eigenvalues of the associated Laplacians \cye{and vice versa}. 
\cye{This eigenvalues optics, upon employing a suitable numerical simulation} over a general triangulation \cye{and some clever post-processing estimates, enables to numerically compute guaranteed upper bounds on the Poincar\'e--Friedrichs constants, see~\cite{Cars_Ged_LB_eigs_14,Liu_fram_eigs_15} for~\eqref{math:pot_PF_grad} and~\cite{gallistl2023computational} for \eqref{math:pot_PF_curl}}. \cye{Our approaches, in contrast, are either analytical or combinatorial, and in any case we only employ the given geometry but no computer simulation.}

@@

\cye{We devote a considerable effort to convex domains for two reasons: first, they are evidently of independent interest, and second, we need them as a building block for the above-discussed recursive procedures on general (non-convex and non-star-shaped) triangulated domains. Here,} we follow closely the \cye{concepts of} Costabel and McIntosh~\cite{costabel2010bogovskiui} who study regularized Poincar\'e and Bogovski\u{\i}-type potential operators. In comparison to their exposition, \cye{which allows for} domains star-shaped with respect to a ball, the regularized potential operators we propose are simplified: we only study them over convex domains and use simpler (constant) weight functions. 
The resulting operator features generally lower regularity \cye{but} its properties are sufficient for our purposes. \cye{Crucially, this allows us to} easily estimate their operator norms.

@@

\cye{We treat specifically the gradient operator (the setting of Section~\ref{section:intro_grad}) on triangulated domains that can be non-convex and non-star-shaped but where the triangulations are face connected (though not necessarily shellable). Here, Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} are our main results.} 
\mwl{MV: remark from after Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} moved here.} 
The computable Poincar\'e--Friedrichs constants obtained here depend on only a few parameters of the \cye{given} triangulation: the length of any traversal from the root simplex, the ratios of the volumes of \cye{the} adjacent simplices, the Poincar\'e--Friedrichs constants on each face patch, and the Poincar\'e--Friedrichs constant of the initial simplex. \cye{Here,} the \cye{face patch} Poincar\'e--Friedrichs constants \cye{are} estimated in terms of shape \cye{regularity parameters of the triangulation}; if the face patches of the triangulation are convex, then better estimates are possible.

\cye{Our main result for the curl and divergence operators (Section~\ref{section:intro_curl_div}) and more generally in the exterior calculus setting (Section~\ref{section:intro_EC}) for triangulated domains are Theorems~\ref{theorem:poincarefriedrichsestimate:exterior} and~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.
They treat triangulated domains that can be non-convex and non-star-shaped but where the triangulation is shellable. As discussed above, though this is slightly less generic than merely face-connected triangulations, this namely covers patches of simplices sharing a given vertex, edge, face, or $n$-simplex (stars) in two and three space dimensions, since these local patches are shellable. The dependencies of the obtained computable Poincar\'e--Friedrichs constants are similar to those of the gradient operator in Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei}.}

@@

    Without loss of generality, $F$ lies in the hyperplane plane of the second to last coordinates,
    so that the height of $T_1$ and $T_2$ is in the first coordinate direction. 
    After further rotation, we can assume that $z_1$ and $z_2$ are in the span of the first three coordinate directions with 
    We can write 
    \begin{align*}
        z_1 = a_1 e_1 + a_2 e_2, \qquad z_2 = b_1 e_1 + b_2 e_2 + e_3 b_3.
    \end{align*}
    The face $F$ spans a hyperplane over which $\Xi$ is the identity, and it remains to study the matrix
    \begin{align*}
        A = 
        \begin{pmatrix}
            b_1/a_1 & 0 & 0
            \\
            (b_2-a_2)/a_1 & 1 & 0
            \\
            b_3/a_1 & 0 & 1
        \end{pmatrix}
        =:
        \begin{pmatrix}
            c_1 & 0 & 0
            \\
            c_2 & 1 & 0
            \\
            c_3 & 0 & 1
        \end{pmatrix}
    \end{align*}
    Its singular values are $1$ and 
    \begin{align*}
        \sigma_{\max}
        =
        \frac{1}{\sqrt 2}
        \sqrt{
            1 + c_1^2 + c_2^2 + c_3^2 
            +
            \sqrt{ ( 1 + c_1^2 + c_2^2 + c_3^2 )^2 - 4 c_1^2 }
        }
        =
        \frac{1}{\sqrt 2}
        \sqrt{
            1 + c_1^2 + c_2^2 + c_3^2 
            +
            \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 + 2 c_1 }
            \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 - 2 c_1 }
        }
        =
        \frac{1}{2}
        \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 + 2 c_1 }
        +
        \frac{1}{2}
        \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 - 2 c_1 }
        =
        \frac{1}{2}
        \sqrt{ ( 1 + c_1 )^2 + c_2^2 + c_3^2 }
        +
        \frac{1}{2}
        \sqrt{ ( 1 - c_1 )^2 + c_2^2 + c_3^2 }
        \leq 
        \sqrt{ ( 1 + c_1 )^2 + c_2^2 + c_3^2 }
        .
    \end{align*}
    \begin{align*}
        \sigma_{\max}
        =
        \frac{1}{\sqrt 2}
        \sqrt{
            1 + c_1^2 + c_2^2 + c_3^2 
            -
            \sqrt{ c_1^2 + c_2^2 + c_3^2 - 4 c_1^2 }
        }
        =
        \frac{1}{2}
        \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 + 2 c_1 }
        -
        \frac{1}{2}
        \sqrt{ 1 + c_1^2 + c_2^2 + c_3^2 - 2 c_1 }
        .
    \end{align*}

@@

\begin{remark}
    % The main feature in proving these results is that the best constant in the Poincar\'e inequality is already the constant in the $1$-dimensional case. 
    Any estimate for the Poincar\'e--Friedrichs constant implies an estimate for the Poincar\'e constant via~\eqref{math:p_vs_pf}. 
    Let us compare the estimate $C_{{\rm P},{\rm CW},p}$ for the Poincar\'e constant
    with the estimate obtained from the Poincar\'e--Friedrichs constant $C_{\PF,{\rm EFNT},p}$. 
    In the case $2 \leq p$, 
    \begin{align*}
        2^{1-\frac 2 p}
        C_{\PF,{\rm EFNT},p}
        =
        \frac{ 2^{1-\frac 2 p} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq 
        4^{-\frac 1 p}
        \leq
        C_{{\rm P},{\rm CW},p} 
        .
    \end{align*}
    In the case $2 \geq p$, 
    \begin{align*}
        2^{\frac 2 p - 1}
        C_{\PF,{\rm EFNT},p}
        =
        \frac{ 2^{\frac 2 p - 1} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq
        \frac{ 2^{\frac 2 p-1} }{2}
        &
        =
        2^{\frac 2 p - 2}
        \\&
        =
        4^{\frac 1 p - 1}
        \leq
        C_{{\rm P},{\rm CW},p} 
%         % 
%         .
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 2\pi \sqrt[p]{p-1} \cdot 2 }
%         =
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 4\pi \sqrt[p]{p-1} }
%         \leq 
%         \frac{1}{2}
        .
    \end{align*}
    It follows that \eqref{math:esposito:poincare} is generally a tighter estimate than \eqref{math:chuawheeden} for $1 < p < \infty$.
    % DONE: what exactly is the limit as p -> 1 ? It converges to 1.
    % DONT: insert the original inequality by chua and Wheeden
    % DONE: chua Wheeden upper bound is not optimal!
    % DONE: Theorem 1.2 in Chua Wheeden 
% DONT: Keep ?
%     In the limit as $p$ goes to $1$, the factor $C_{\PF,{\rm EFNT},p}$ converges to $1/2$,
%     which is the optimal Poincar\'e--Friedrichs constant and also the optimal factor in the Poincar\'e inequality \eqref{math:acostaduran}
%     \begin{align*}
%         \| u - u_{\Omega} \|_{L^{1}(\Omega)}
%         \leq 
%         \frac{1}{2}
%         \diam(\Omega)
%         \| \nabla u \|_{L^{1}(\Omega)}
%         ,
%         \quad
%         \forall 
%         u \in W^{1,1}(\Omega)
%         .
%     \end{align*}
%     In the limit as $p$ goes to infinity, the factor $C_{\PF,{\rm EFNT},p}$ converges to $1/2$, 
%     which implies the Poincar\'e inequality 
%     \begin{align*}
%         \| u - u_{\Omega} \|_{L^{\infty}(\Omega)}
%         \leq 
%         \diam(\Omega)
%         \| \nabla u \|_{L^{\infty}(\Omega)}
%         ,
%         \quad
%         \forall 
%         u \in W^{1,\infty}(\Omega)
%         %
%     \end{align*}
%     because the $L^{\infty}$-norm is the limit of the $L^{p}$-norm. % DONT: cite 
%     % DONT: review this entire result 
\end{remark}

@@

\begin{remark}
    % The main feature in proving these results is that the best constant in the Poincar\'e inequality is already the constant in the $1$-dimensional case. 
    Any estimate for the Poincar\'e--Friedrichs constant implies an estimate for the Poincar\'e constant, via~\eqref{math:p_vs_pf}. 
    Let us compare the estimate $C_{{\rm P},{\rm CW},p}$ for the Poincar\'e constant
    with the estimate obtained from the Poincar\'e--Friedrichs constant $C_{\PF,{\rm EFNT},p}$. 
    In the case $2 \leq p$, 
    \begin{align*}
        2^{1-\frac 2 p}
        C_{\PF,{\rm EFNT},p}
        =
        \frac{ 2^{1-\frac 2 p} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq 
        4^{-\frac 1 p}
        \leq
        C_{{\rm P},{\rm CW},p} 
        .
    \end{align*}
    In the case $2 \geq p$, 
    \begin{align*}
        2^{\frac 2 p - 1}
        C_{\PF,{\rm EFNT},p}
        =
        \frac{ 2^{\frac 2 p - 1} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq
        \frac{ 2^{\frac 2 p-1} }{2}
        &
        =
        2^{\frac 2 p - 2}
        \\&
        =
        4^{\frac 1 p - 1}
        \leq
        C_{{\rm P},{\rm CW},p} 
%         % 
%         .
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 2\pi \sqrt[p]{p-1} \cdot 2 }
%         =
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 4\pi \sqrt[p]{p-1} }
%         \leq 
%         \frac{1}{2}
        .
    \end{align*}
    It follows that \eqref{math:esposito:poincare} is generally a tighter estimate than \eqref{math:chuawheeden} for $1 < p < \infty$.
    % DONE: what exactly is the limit as p -> 1 ? It converges to 1.
    % DONT: insert the original inequality by chua and Wheeden
    % DONE: chua Wheeden upper bound is not optimal!
    % DONE: Theorem 1.2 in Chua Wheeden 
% DONT: Keep ?
%     In the limit as $p$ goes to $1$, the factor $C_{\PF,{\rm EFNT},p}$ converges to $1/2$,
%     which is the optimal Poincar\'e--Friedrichs constant and also the optimal factor in the Poincar\'e inequality \eqref{math:acostaduran}
%     \begin{align*}
%         \| u - u_{\Omega} \|_{L^{1}(\Omega)}
%         \leq 
%         \frac{1}{2}
%         \diam(\Omega)
%         \| \nabla u \|_{L^{1}(\Omega)}
%         ,
%         \quad
%         \forall 
%         u \in W^{1,1}(\Omega)
%         .
%     \end{align*}
%     In the limit as $p$ goes to infinity, the factor $C_{\PF,{\rm EFNT},p}$ converges to $1/2$, 
%     which implies the Poincar\'e inequality 
%     \begin{align*}
%         \| u - u_{\Omega} \|_{L^{\infty}(\Omega)}
%         \leq 
%         \diam(\Omega)
%         \| \nabla u \|_{L^{\infty}(\Omega)}
%         ,
%         \quad
%         \forall 
%         u \in W^{1,\infty}(\Omega)
%         %
%     \end{align*}
%     because the $L^{\infty}$-norm is the limit of the $L^{p}$-norm. % DONT: cite 
%     % DONT: review this entire result 
\end{remark}

@@

% While the underlying idea remains the same, they differ in the way the intermediate potentials are extended.
% This is described in Theorem~\ref{theorem:poincarefriedrichsestimate:grad} and relies on Lemma~\ref{lemma:poincarefriedrichsoverfacepatch} as an auxiliary result. 
% This is described in Theorem~\ref{theorem:poincarefriedrichsestimate:gradzwei}.

@@

\begin{remark}
    The Poincar\'e--Friedrichs constant obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad} increasingly overestimates the best one as number of $n$-simplices in the triangulation $\calT$ increases. We hence target its use to local patches (stars), in particular non-convex boundary stars. The latter occur inevitably at reentrant corners. Clearly, the same building principle applies whenever we have any non-overlapping partition of $\overline \Omega$ into convex local patches of $n$-simplices $\{ U_m \}$ with a face-connected path. The proof proceeds verbatim, where we merely replace the simplices $\{ T_{m} \}$ by the convex local patches $\{ U_m \}$. This may allow to a large decrease of the number of elements in the partition of $\overline \Omega$ and thus to a largely improved estimate of the best Poincar\'e--Friedrichs constant. 
\end{remark}

@@

We now collect important examples of shellable triangulations. Essentially, in two space dimensions, interesting triangulations shellable, but starting from three space dimensions, non-shellable situations can arise. Our main interest are local patches (stars) within triangulations: these are shellable up to three space dimensions, but not necessarily beyond.
